---
title: "Practice 3"
output: html_notebook
---

```{r}
library(matlib)
library(rlang)
library(openxlsx)
library(Seurat)
library(Matrix)
library(dplyr)
```

## Using the raw data in file “SCOplanaria.txt” do:
1. Initialize a Seurat object.
2. Perform quality control, normalization, variable features selection and scaling of the data.
3. Perform a PCA analysis and obtain a t-SNE clustering plot.

```{r}
library(patchwork)
# Uploading data
data <- read.table("SCOplanaria.txt", row.names=1) 
```

## My own parameters
```{r}
# Create a new Seurat Object --------------------------------------------
dataMatrix = as.matrix(data)                               # For this task it is mandatory to create a matrix
datasparseMatrix = Matrix(data=dataMatrix, sparse=TRUE)    # And sparse it

# Using my own parametres:
seurat1 = CreateSeuratObject(counts=datasparseMatrix, project="SCOP", min.cells=1, min.features=1)

#Quality Control --------------------------------------------------------
head(seurat1@meta.data, 5)
# Visualize QC metrics as a violin plot
VlnPlot(seurat1, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)


# Normalizing the data -------------------------------------------------
seurat1[["RNA"]]@data
seurat1<- NormalizeData(seurat1)
Normalized.Reads <- colSums(seurat1[["RNA"]]@data)
seurat1 <- AddMetaData (seurat1,Normalized.Reads,col.name="Normalized_nCount_RNA")
#print(Normalized.Reads)
#head(seurat1[["RNA"]]@data, 5)
#head(seurat1@meta.data, 5)
VlnPlot(object = seurat1, features= c("nCount_RNA", "Normalized_nCount_RNA"), ncol = 2)


# Feature selection --------------------------------------------------
seurat1 <- FindVariableFeatures(seurat1, selection.method = "vst", nfeatures = 1000)
# plot variable features without labels
plot1 <-  LabelPoints(plot=VariableFeaturePlot(seurat1), 
                 points=head(VariableFeatures(seurat1),3), repel=TRUE)
plot1

# Scaling the data --------------------------------------------------
all.genes <- rownames(seurat1)
seurat1 <- ScaleData(seurat1, features = all.genes)

# Perform a PCA analysis -------------------------------------------
seurat1pca <- RunPCA(seurat1, features = VariableFeatures(object = seurat1))

# Obtain a t-SNE clustering plot -----------------------------------
seurat1pcaplot = FindNeighbors(seurat1pca, dims=1:5)
seurat1pcaplot = FindClusters(seurat1pcaplot, resolution=0.2)
seurat1 <- RunTSNE(seurat1pcaplot, dims = 1:5)
DimPlot(seurat1, reduction = "tsne")

```
## Given paramters

```{r}
# Create a new Seurat Object
dataMatrix = as.matrix(data)                               # For this task it is mandatory to create a matrix
datasparseMatrix = Matrix(data=dataMatrix, sparse=TRUE)    # And sparse it

# Inlcuding new parametres:
#Keep genes expressed in at least 3 cells and cells with at least 200 genes detected.
seurat2 = CreateSeuratObject(counts=datasparseMatrix, project="SCOP", min.cells=3, min.features=200)

#Quality Control is automatically calculated during CreateSeuratObject()
head(seurat2@meta.data, 5)
# Visualize QC metrics as a violin plot
VlnPlot(seurat2, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)

# Keep cells with 200- 2500 number of detected genes
seurat2Subset = subset(seurat2, subset=nFeature_RNA>200&nFeature_RNA<2500)

# Use LogNormalize method with a scale factor of 10000.
seurat2Subset[["RNA"]]@data
seurat2Normalizated = NormalizeData(seurat2Subset, normalization.method="LogNormalize", scale.factor=10000)

Normalized.nCounts <- colSums(seurat2Normalizated[["RNA"]]@data)
seurat2Normalizated <- AddMetaData (seurat2Normalizated,Normalized.nCounts,col.name="Normalized_nCount_RNA")
VlnPlot(object = seurat2Normalizated, features= c("nCount_RNA", "Normalized_nCount_RNA"), ncol = 2)

# Select 3000 variable features
seurat2Normalizated <- FindVariableFeatures(seurat2Normalizated, selection.method = "vst", nfeatures = 3000)
plot2 = LabelPoints(plot=VariableFeaturePlot(seurat2Normalizated), 
                 points=head(VariableFeatures(seurat2Normalizated),2), repel=TRUE)
plot2

# Regress out variability for number of genes in each cell.
seurat2Normalizated = ScaleData(seurat2Normalizated, features=rownames(seurat2Normalizated))

# Use PCs 1-5 and a resolution value of 0.6, t-SNE 
seurat2_pca <- RunPCA(seurat2Normalizated, features = VariableFeatures(object = seurat2Normalizated))

seurat2_pca_plot = FindNeighbors(seurat2_pca, dims=1:5)
seurat2_pca_plot = FindClusters(seurat2_pca_plot, resolution=0.6)
seurat2 <- RunTSNE(seurat2_pca_plot, dims = 1:5)
DimPlot(seurat2, reduction = "tsne")
```

## Extract a table of the 5 top biomarkers for each of your clusters. Use these biomarkers to generate a heatmap. Comment the results

```{r}
biomarkers = FindAllMarkers(seurat2, only.pos=TRUE, min.pct=0.25, logfc.threshold=0.25)
biomarkers_top5 = biomarkers %>% group_by(cluster) %>% top_n(n=5)

#Visualize as heatmap
DoHeatmap(seurat2, features=biomarkers_top5$gene) 
```


## Using VlnPlot() and FeaturePlot() functions, show the expression distribution of these markers along your clusters. With this information, identify and rename your clusters. 
```{r}
cluster.id = c("Early epidermal progenitors", "Late epidermal progenitors",
                  "Epidermis", "Muscle progenitors",
                  "Muscle body", "Neural progenitors",
                  "GABA neurons", "Phagocytes",
                  "Parenchymal cells", "Pigment")

names(cluster.id) <- levels(seurat2)
seurat2 <- RenameIdents(seurat2, cluster.id)
DimPlot(seurat2, reduction = "tsne", label = TRUE, pt.size = 0.5) + NoLegend()

VlnPlot(seurat2, features= "dd-Smed-v6-61-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-2178-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-298-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-1410-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-702-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-2548-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-9977-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-48-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-175-0")   
VlnPlot(seurat2, features= "dd-Smed-v6-1161-1")   

FeaturePlot(seurat2, features= "dd-Smed-v6-61-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-2178-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-298-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-1410-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-702-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-2548-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-9977-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-48-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-175-0")   
FeaturePlot(seurat2, features= "dd-Smed-v6-1161-1")


```
## Compare your results with this lineage tree reconstruction of planarian cell types (Plass et al., 2018). What cell type do you think cluster 0 (the central cluster) 
```{r}
new.cluster.ids <- c("Neoblast", "neural progenitors",
                     "early epidermal progenitors", "epidermis", 
                     "GABA neurons", "parenchymal cells",
                     "muscle body", "late epidermal progenitors",
                     "mucle progenitors", "pigment", 
                     "phagocytes")

names(new.cluster.ids) <- levels(seurat2)
seurat2 <- RenameIdents(seurat2, new.cluster.ids)
DimPlot(seurat2, reduction = "tsne", label = TRUE, pt.size = 0.5) + NoLegend()
```

# dd-Smed-v6-1999-0 is a neoblast (stem) marker gene. Show its expression distribution with VlnPlot() and FeaturePlot() and explain the result.
```{r}
VlnPlot(seurat2, features = "dd-Smed-v6-1999-0")                  
FeaturePlot(seurat2, features = "dd-Smed-v6-1999-0")
DimPlot(seurat2, reduction = "tsne", label = TRUE, pt.size = 0.5) + NoLegend()

```

