---
title: "Practice 1"
output: html_notebook
---

## Calculate PCA among ground tissue samples for shr mutant, the wild type andthe complemented lines with several transcription factor
```{r}
# Uploading data
tablecsv <-read.csv("table_ex_1.csv",row.names=1)

# Print data characteristics
print("Column names")
colnames(tablecsv)  # Column names

print("Data size")
dim(tablecsv)       # Data size

print("Data saved as:")
class(tablecsv)     # Data class (data frame)

```

```{r}
# Chose first 8 columns (needed information)
data <- tablecsv[,1:8]

# Keep gene id as rows
rownames(data) <-rownames(tablecsv)

# Change column names to be more clear
colnames(data) <- c("shrJ0571","WT","shrJ0571+BLJ","shrJ0571+JKD","shrJ0571+MGP","shrJ0571+NUC","shrJ0571+IME","shrJ0571+SCR")

# Show the result
print(data)
```

```{r}
# Perform the PCA with the data prepared on last section
PCA<-princomp(data,cor=FALSE,scores=TRUE)
plot(PCA) # Print PCA results as a plot

#Print PCA loaings and plot them
plot(PCA$loadings)
text(PCA$loadings,labels=row.names(PCA$loadings),pos=c(2,4,4,4,2,4)) # pos is the position of the label (1 = up, 2 = left, 3 = down and 4 = right)
PCA$loadings 

```

## Create intermediate transcriptomes between shr mutant and the wild type (J0571)which represent 25%, 50% and 75% of recovery (complementation) in gene expres-sion.

```{r}
# Recalculate each transcriptomic expression
data2 <- transform(data, Complementation_25=(0.25*(WT)+0.75*shrJ0571), Complementation_50=(0.50*(WT)+0.50*shrJ0571), Complementation_75=(0.75*(WT)+0.25*shrJ0571))

# Perform a PCA and plot it
PCA2<-princomp(data2, cor=FALSE,scores=TRUE)
plot(PCA2)

#Print PCA loaings and plot them
plot(PCA2$loadings)
text(PCA2$loadings,labels=row.names(PCA2$loadings),pos=c(2,4,4,4,2,4,2,4,2))

# Fast nterpretation
# WT is Complementation_75
# MGP Y IME are Complementation_50
# BLJ is Complementation_25

# With this study we will know for each mutant with trasncription factor if it is near the WT or the mutant and in which percentaje

```
## Add the transcriptome of cells corresponding to SCR domain and recalculate PCA. Plot variance for components and loadings.

```{r}
# Join SCRdomain column of the given data to the data we are studying
data3<-cbind(data2, tablecsv$SCRdomain )
# Add a new column name for it
colnames(data3) <- c("shrJ0571","WT","shrJ0571+BLJ","shrJ0571+JKD","shrJ0571+MGP","shrJ0571+NUC","shrJ0571+IME","shrJ0571+SCR", "Complementation_25", "Complementation_50", "Complementation_75","SCRdomain")
# Keep genes id as row names
rownames(data3) <-rownames(tablecsv)

# Perform a new PCA and plot it
PCA3<-princomp(data3,cor=FALSE,scores=TRUE)
plot(PCA3)

# Print and show PCA loadings
plot(PCA3$loadings)
text(PCA3$loadings,labels=row.names(PCA3$loadings),pos=c(2,2,4,4,2,4,4,4,2))
PCA3$loadings 

```


## Find the most important genes which contributing to observed transcriptomic changes by extracting genes (~20) with highest and lowest score values
```{r}

# Yelow will be the top expression
newcolors<-colorRampPalette(colors=c("gray48","springgreen3","yellow"))(256)

# Sort PCA scores odf component 1 (most expressed first) and take:
sorted_comp1<-sort(PCA3$scores[,1])
top20_comp1<-head(sorted_comp1, 20)   # The 20 first (more expressed)
less20_comp1<-tail(sorted_comp1,20)   # And the 20 last (less expressed)

# Sort PCA scores odf component 1 (most expressed first) and take:
sorted_comp2<-sort(PCA3$scores[,2])
top20_comp2<-head(sorted_comp2, 20)   # The 20 first (more expressed)
less20_comp2<-tail(sorted_comp2,20)   # And the 20 last (less expressed)

# Join top_20
top_20_all=c(top20_comp1,top20_comp2)
top_20_matrix=as.matrix(top_20_all)
top20=data3[rownames(top_20_matrix),]
top_20_final=as.matrix(top20)
heatmap(top_20_final, col = newcolors)

# Join less_20
less_20_all=c(less20_comp1,less20_comp2)
less_20_matrix=as.matrix(less_20_all)
less20=data3[rownames(less_20_matrix),]
less_20_final=as.matrix(less20)
heatmap(less_20_final, col = newcolors)

```

